Dark matter halo shape at $z=0$ in the Auriga simulations: radial evolution and alignment with the stellar disk

### 2019-07-19 Document/paper.tex
edit

% Abstract of the paper
\begin{abstract}
We present shape measurements of dark matter halos at redshift $z=0$
in a suite of 30 cosmological simulations from the Auriga project. 
We compare the results in full magnetohydrodynamics against dark
matter only physics. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
This change in triaxility is more pronounced towards the inner
regions of the halo.
In simulations with baryons we measure a twist in the dark matter
shape as a function of radial distance; this effect is almost absent
in dark matter only simulations.
We quantify this twisting effect as a changing degree of alignment
bewteen the halo shape with the stellar disk shape as a function of radius.
An almost perfect alignment between the two components appears at
$0.25R_{200}\sim56$ kpc.  
We show that the twist in the radial shape evolution correlates
positively with triaxility.
Furthermore, at distances $\lesssim 30$ kpc, rounder and coherent dark
matter distributions correlate with extended massive stellar disks and
low gas densities. 
In a comparison against observational constraints of dark matter halo
triaxiality for the Milky Way we find that $20\%$ of halos our sample
are consistent with observational results derived from the Pal 5 stream
that favor $b/a\approx 1$ and $c/a>0.77$. Including baryons is a
required element to achieve this level of agreement.
In contrast, all simulations (dark matter only and with baryons) are at
odds with the constrains derived from the Saggitarius stream that
favor $b/a\approx 0.97$ and $c/a\approx0.44$.
\end{abstract}

>>>>

% Abstract of the paper
\begin{abstract}
We present shape measurements of dark matter halos at redshift $z=0$
in a suite of 30 zoom-in simulations from the Auriga project. 
We compare the results in full magnetohydrodynamics against dark
matter only physics. 
We find a strong influence of baryons in making dark matter halos
rounder at all radii compared to its dark matter only counterparts.
This change in triaxility is more pronounced towards the inner
regions of the halo.
At distances $\lesssim 30$ kpc, rounder and coherent dark matter
distributions correlate with extended massive stellar disks and low gas
densities.  
We measure the alignment between the halo and the stellar disk and
find almost perfect alignment at $0.25R_{200}\sim56$ kpc.   
The alignment between the two components changes as a function of
radius implying that the halo shape twists; this effect is almost absent in the dark
matter only simulations. 
We quantify the twisting effect wtih the standard deviation of the
angle between the minor axis of the disk and the halo for radii
$\leq R_{200}$ to show that the twisting correlates positively with
extended stellar disks and low gas densities.
In a comparison against observational constraints of dark matter halo
triaxiality for the Milky Way we find that $20\%$ of halos our sample
are consistent with observational results derived from the Pal 5 stream
that favor $b/a\approx 1$ and $c/a>0.77$. Including baryons is a
required element to achieve this level of agreement.
In contrast, all simulations (dark matter only and with baryons) are at
odds with the constrains derived from the Saggitarius stream that
favor $b/a\approx 0.97$ and $c/a\approx0.44$.
\end{abstract}
